The \project{unRAVE} catalog

### 2016-08-23 article/unrave.tex
Expanded the stubs in the introduction

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\rave}{\project{\acronym{RAVE}}}

@@

@@

\begin{abstract}
The Milky Way may be our greatest laboratory for understanding galaxy formation.
This premise relies on the ability to measure [1] 3D positions and motions, and
[2] detailed chemical abundances, for large numbers of individual stars.  On 14
September 2016 the \gaia\ mission will release the former for $\approx2\times10^6$
stars.  However, a catalog of precise chemical abundances to complement that data
release is currently lacking.  The RAdial Velocity Experiment (\rave) survey has 
acquired spectra for 292,036 stars in common with the first \gaia\ data release, 
constituting the largest overlap of any stellar spectroscopic sample.  Here we 
perform an independent analysis of \rave\ spectra using an implementation of 
\thecannon\ that uses prior probabilities on atomic line formation and stellar
astrophysical parameters.  Our approach yields precise stellar labels at lower
S/N ratios than most \emph{ab initio} approaches, allowing us to deliver
improved radial velocities, effective temperature $\teff$, surface gravity $\logg$,
and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) 
for \Nstars\ stars.  We validate our results internally from repeat visits, 
and externally against existing high-resolution spectroscopic surveys.  The 
typical precision in chemical abundances is X.XX~dex.  We provide 
spectrophotometric distances based on our labels, improving upon the parallax
distances available for the fainter targets in the first \gaia\ data release.
\stub{A one-line science result based on RAVE/Hipparcos overlap}
\end{abstract}

@@

The Milky Way is considered to be the best laboratory for understanding galaxy
formation and evolution.  This premise hinges on the ability to precisely measure 
the position, motion, and chemistry of (many) individual stars, and use those data 
to infer the structure, kinematics, and chemical enrichment of the Galaxy 
\citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,
Casey_2014a,Casey_2014b}.  % include references on: disk chemistry, halo structures
However, these quantities are not known for even 1\% of stars in the Milky Way.  
Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics
from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
stellar spectroscopists frequently report significantly different chemical 
abundance patterns \citep{Smiljanic_2014}.  Countless long-standing astrophysics
problems could be resolved if those measurements were known with high fidelity: 
\stub{x? \citep{someone}; y; and z, to list just a few}.

\stub{Introducing Gaia; rely on ground-based spectroscopic surveys for chemistry}
% Introducing Gaia
% Gaia will produce a very clear snapshot, producing astrometry space motions
% What will Gaia provide eventually.
% In the first data release what do we get.
% While this is an order of magnitude better than Hipparcos, if we want to use these data to understand the chemical evolution of the galaxy then we need additional information.
% until abundances are available from gaia RVS, we rely on ground-based spectroscopic surveys to provide the chemistry

@@

\section{Model}
\label{sec:model}

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni. We note that the while Ti can be measured, its values in both 
APOGEE DR12 and DR13 are known to not follow the expected Galactic trends \citep{Holtzman_2015, Hawkins_2016}.
This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars sit above the isochrones, not below).  If we move to lower-resolution studies,
there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
seen on the main-sequence: the labels are expectedly less precise given the lower
resolution, no detailed abundance information is available\footnote{Abundance 
information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
sample contains only giant stars}, and there is only sparse sampling across the 
Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
but results from that sample are not yet precise enough for our purposes.

@@

Assembling a training set for the main-sequence and sub-giant branch was more complex.
% training set from K2/EPIC?!!
% Identifying obviously bad stars that have been mis-classified by K2/EPIC

% Magic wizardry for missing and imperfect labels.

@@

% Censoring masks on abundance labels
% Priors on astrophysical parameters from isochrones

@@

\stub{What abundances are available?}
\stub{Any differences in RAVE abundances?}

@@

@@

@@

\begin{figure}[p]
\caption{A H-R diagram showing the test set labels.\label{fig:test-set-hrd}}
\end{figure}

>>>>

@@

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronymn{APOGEE}}}
\newcommand{\lamost}{\project{\acronymn{LAMOST}}}
\newcommand{\hipparcos}{\project{Hipparcos}}

@@

\begin{abstract}
The Milky Way may be our greatest laboratory for understanding galaxy formation.
This premise relies on the ability to measure --- for large numbers of stars ---
3D positions and motions, and their detailed chemical abundances.  On 14 September
2016 the \gaia\ mission will release the former for $\approx2\times10^6$ stars.
However, a catalog of precise chemical abundances to complement that data
release is currently lacking.  The RAdial Velocity Experiment (\rave) survey has 
acquired spectra for 292,036 stars in common with the first \gaia\ data release, 
constituting the largest overlap of any stellar spectroscopic sample.  Here we 
perform an independent analysis of \rave\ spectra using an implementation of 
\thecannon\ that uses prior probabilities on atomic line formation and stellar
astrophysical parameters.  Our approach yields precise stellar labels at lower
S/N ratios than most \emph{ab initio} approaches, allowing us to deliver
improved radial velocities, effective temperature $\teff$, surface gravity $\logg$,
and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) 
for \Nstars\ stars.  We validate our results internally from repeat visits, 
and externally against existing high-resolution spectroscopic surveys.  The 
typical precision in chemical abundances is X.XX~dex.  We provide 
spectrophotometric distances based on our labels, improving upon the parallax
distances available for the fainter targets in the first \gaia\ data release.
\stub{A one-line science result based on RAVE/Hipparcos overlap}
\end{abstract}

@@

The Milky Way is considered to be the best laboratory for understanding galaxy
formation and evolution.  This premise hinges on the ability to precisely measure 
the astrometry and chemistry for (many) individual stars, and to use those data 
to infer the structure, kinematics, and chemical enrichment of the Galaxy 
\citep[e.g.,][]{Schlaufman_2009,Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,
Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Bovy_2016}.  However, these 
quantities are not known for even 1\% of stars in the Milky Way.  Stellar 
distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,
Madler_2016}, proper motions can be plagued by unquantified systematics from 
the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
stellar spectroscopists frequently report significantly different chemical 
abundance patterns \citep{Smiljanic_2014}.  The impact this lack of data has on
scientific inferences cannot be understated.  It limits our understanding on a
number of sub-fields in astrophysics, including the properties of exoplanet host
stars, the formation (and destruction) of stars and clusters, as well as the 
structure of the Galactic disk, to name a few.

These formed just some of the scientific drivers for the \gaia\ space telescope.
\gaia\ is primarily an astrometric mission, providing precise positions,
parallaxes and proper motions for more than $10^9$ stars in its final data
release, scheduled for 2022.  While this is a sample size about three orders 
of magnitude larger than its predecessor \hipparcos, both astrometry and 
chemistry are required to fully characterize the formation and evolution of our Galaxy.  
For this reason \gaia\ will also provide radial velocities, stellar parameters 
and chemical abundances for a subset of brighter stars, primarily from the on-board
radial velocity spectrometer.  However, those chemical abundances will not be 
available until later releases.  Until those abundances are available,
astronomers seeking to use chemistry and kinematic measurements are reliant on 
ground-based spectroscopic surveys in order to complement the available astrometry
from \gaia.

@@

\section{Method}
\label{sec:method}

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars sit above the isochrones, not below).  If we move to lower-resolution studies,
there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
seen on the main-sequence: the labels are expectedly less precise given the lower
resolution, no detailed abundance information is available\footnote{Abundance 
information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
sample contains only giant stars}, and there is only sparse sampling across the 
Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
but results from that sample are not yet precise enough for our purposes.

@@

Assembling a training set for the main-sequence and sub-giant branch was more complex.
% training set from K2/EPIC?!!
% Identifying obviously bad stars that have been mis-classified by K2/EPIC
% Refer to the training set figure.
% Two separate models, with an overlap region.

% Initially we train on three labels.

@@

% Magic wizardry for missing and imperfect labels.
% Censoring masks on abundance labels
% Priors on astrophysical parameters from isochrones

@@

\stub{What abundances are available?}

% This is more relevant for the discussion, not description of the training set.
We note that the while Ti can be measured, its values in both 
APOGEE DR12 and DR13 are known to not follow the expected Galactic trends \citep{Holtzman_2015, Hawkins_2016}.

\stub{Any differences in RAVE abundances?}

@@

\bibitem[Bovy et al.(2016)]{Bovy_2016} Bovy, J., Rix, H.-W., Schlafly, E.~F., et al.\ 2016, \apj, 823, 30

@@

\bibitem[Schlaufman et al.(2009)]{Schlaufman_2009} Schlaufman, K.~C., Rockosi, C.~M., Allende Prieto, C., et al.\ 2009, \apj, 703, 2177

@@

\begin{figure}[p]
\includegraphics[width=\textwidth]{figures/hrd-test-set.pdf}
\caption{A `spectroscopic' Hertzsprung-Russell diagram showing the effective temperature $\teff$ and surface gravity $\logg$ for stars in the test set.  Only stars with $\chi_r < 3$ are shown, and each panel shows stars that meet a threshold in $S/N$ ratio.\label{fig:test-set-hrd}}
\end{figure}
